---
title: "Prediction Assignment"
author: "swathy"
date: "3/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.width=10, fig.height=5)
options(width=120)
```

## The Project


The goal of this project is to predict the manner in which they did the exercise.

## The Data

The training data for this project are available here:

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

The test data are available here:

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>



## Loading the Data & Libraries needed into R:

```{r, echo=TRUE}
library(lattice)
library(ggplot2)
library(plyr)
library(randomForest)
```

```{r, echo=TRUE}
# Download the training data
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
              destfile = "./pml-training.csv")
# Load the training dataset
dt_training <- read.csv("./pml-training.csv", na.strings=c("NA","#DIV/0!",""))
# Download the testing data
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "./pml-testing.csv")
# Load the testing dataset
dt_testing <- read.csv("./pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

## Cleaning the Data

1. Remove all columns that contains NA and remove features that are not in the testing dataset.
2. The testing dataset has no time-dependence, therefore these values are useless and can be disregarded.
3. Also remove the first 7 features since they are related to the time-series or are not numeric.

```{r, echo=TRUE}
features <- names(dt_testing[,colSums(is.na(dt_testing)) == 0])[8:59]
# Only use features used in testing cases.
dt_training <- dt_training[,c(features,"classe")]
dt_testing <- dt_testing[,c(features,"problem_id")]
dim(dt_training); dim(dt_testing);
```
## Exploratory Data Analysis

1. Split data into a training data set (60% of the total cases) and a testing data set (40% of the total cases). This will allow us to estimate the out of sample error of our predictor.

```{r, echo=TRUE}
set.seed(12345)
library(caret)
inTrain <- createDataPartition(dt_training$classe, p=0.6, list=FALSE)
training <- dt_training[inTrain,]
testing <- dt_training[-inTrain,]
dim(training); dim(testing);
```

## Model Selection: Decision Tree Model

```{r, echo=TRUE}
set.seed(12345)
library(rpart)
  
```

```{r, echo=TRUE}
set.seed(12345)
library(caret)


library(caret)
library(splines)
library(parallel)
library(gbm)
library(randomForest)
modFitRF <- randomForest(classe ~ ., data = training, method = "rf", importance = T, trControl = trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10))
plot(modFitRF)
```

## Building the Boosting Model

```{r, echo=TRUE}
modFitBoost <- train(classe ~ ., method = "gbm", data = training,
                    verbose = F,
                    trControl = trainControl(method = "cv", number = 10))
modFitBoost
plot(modFitBoost)
```

## Predicting with the Random Forest Model

```{r, echo=TRUE}
prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```

## Predicting with the Boosting Model

```{r, echo=TRUE}
prediction <- predict(modFitBoost, testing)
confusionMatrix(prediction, testing$classe)
```

### Predicting with the Testing Data
```

## Random Forest Prediction

```{r, echo=TRUE}
predictionRF <- predict(modFitRF, dt_testing)
predictionRF
```

## Boosting Prediction

```{r, echo=TRUE}
predictionBoost <- predict(modFitBoost, dt_testing)
predictionBoost
```

### Conclusion

```{r, echo=TRUE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predictionRF)
```